Nanophotonics with optical phonons in two dimensions

### 2018-12-21 paper.tex
Edits and comments from Thomas

\usepackage{graphicx}
\usepackage[centering,hmargin=20mm,tmargin=30mm,bmargin=25mm]{geometry}
\usepackage{multirow}
\usepackage{newtxtext}
\usepackage[cmintegrals]{newtxmath}

%----- References -----
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks,
	linkcolor={blue!75!black!80!yellow},
	citecolor={blue!75!black!80!yellow},
	urlcolor={blue!75!black!80!yellow}
}

%----- Captions in sans font -----
\makeatletter
\renewcommand\@make@capt@title[2]{%
	\@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
	\sffamily{\textbf{#1}}\@caption@fignum@sep#2
}%
\renewcommand\figurename{Figure}
\makeatother

@@

\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\sgn}{\operatorname{sgn}}

@@

@@

%\usepackage[usenames]{color}
%\newcommand{\edited}[1]{{\color{red} #1}}
\newcommand{\comment}[2]{\textcolor{blue!70!black}{\small\textsf{\textsuperscript{\textsc{\MakeLowercase{#1}}}[#2]}}}
\usepackage[normalem]{ulem}
\newcommand{\swap}[2]{\textcolor{red!70!black}{\sout{#1}}\textrightarrow\textcolor{green!50!black}{#2}}
\newcommand{\remove}[1]{\textcolor{red!70!black}{\sout{#1}}}
\newcommand{\inset}[1]{\textcolor{green!50!black}{#1}}
\begin{document}

\title{Nanophotonics with optical phonons in two dimensions}

\author{Nicholas Rivera}\email{nrivera@seas.harvard.edu}\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
\author{Thomas Christensen}\affiliation{\MITPhy}
\author{Prineha Narang}\email{prineha@seas.harvard.edu}\affiliation{\HarvardSEAS}

@@

\begin{abstract}
Extreme confinement of electromagnetic energy by phonon-polaritons promises extremely strong and novel forms of control over the dynamics of matter. To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems. Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse (LO and TO) optical phonons is absent at the $\Gamma$ point.  Does this lack of LO-TO splitting imply the lack of a strongly confined electromagnetic (phonon-polariton) mode in polar monolayers? Here, we derive a universal form for the spatially-nonlocal dielectric function of a polar monolayer specified by: the TO phonon frequency, the non-zero group velocity of the LO phonon, and the phonon lifetime. Our analysis reveals that the phonon-polariton of the bulk is ``replaced" by the LO phonon. We present the confinement and propagation losses of LO phonons, finding that high LO phonon confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques. We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to 2D phonon channels, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.  \end{abstract}

@@

Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies. Phonon polaritons share many features in common with plasmon polaritons in conductors. In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}. Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}. In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness. In many cases, a monolayer is the ultimate limit of this effect. Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions. However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point.

In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions. Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected. We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing multi-mode strong coupling between an external emitter and optical phonons in hexagonal boron nitride.

\section{Electrodynamics of optical phonons in two-dimensions}

In this section, we develop the theory of electromagnetic waves associated with optical phonons in two dimensions. The crux of the theory is determining the dielectric permittivity associated with the monolayer. It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$. The coupling Hamiltonian is then given by

@@

H_{\mathrm{int}} = \int d^2x \rho \phi = -\int d^2x (\nabla\cdot\mathbf{P})\phi,

@@

with $\rho$ the charge density and $\mathbf{P}$ the polarization density associated with the ionic motion. The polarization can be expanded in powers of the displacement of the ions. In particular, we can consider the polarization $P_i$ generated in the monolayer when atom $\kappa$ of the unit cell is displaced along the $j$ direction by amount $u_{\kappa,j}$. To lowest order:

@@

where $Z_{\kappa,ij}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area. With this relation between polarization and ionic displacements, we have expressed the Hamiltonian as a coupling between the scalar potential and ionic displacements. We consider the response of the monolayer to a scalar potential of the form $\phi = \phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer. Such a potential corresponds to a longitudinal electric field of the form $\mathbf{E} = i\mathbf{q}\phi(\mathbf{q},\omega)e^{i\mathbf{q}\cdot\mathbf{r}-i\omega t}$. It follows that the induced polarization takes the form $P_i  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,\mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization. The polarization-polarization response function is given by

@@

\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),

@@

where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature. Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by: 

@@

\Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q}L}}{\omega^2_{\mathbf{q}L}-\omega^2}|\hat{q}\cdot\langle 1_{\mathbf{q}L}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2,

@@

where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$. The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!]. This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility. Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response is an uncontrolled double-counting. The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  

@@

To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations. A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer, considering optical phonons in the so-called upper Reststrahlen band which in bulk spans the frequency range of 1360 to 1610 cm$^{-1}$. In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode. In practice, it is the TM mode which is associated with highly confined electromagnetic waves that are of use in nanophotonics. TE waves are not supported at the same frequency as transverse magnetic waves, and require exotic conditions to be realized in two-dimensional materials. Given the isotropy, we may suppress indices from the response tensor. We may also consider without further loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)e^{iqx-i\omega t}$ along the $y$-direction of the monolayer. The direction transverse to the monolayer is denoted as $z$.  With these definitions in place, the Maxwell equation satisfied by the magnetic field is \begin{equation}\label{eq:2dmaxwell}
\left(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \right)H(z) = 0.
\end{equation}
We consider a solution of the form $H(z) = e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = ce^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$. The boundary condition on the magnetic field is $H_y^{(+)}-H_y^{(-)} = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer. This condition enforces $c =(1+\frac{\epsilon_0}{\epsilon_2} \kappa_2\Pi)^{-1}$. Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}c$. These two conditions fully specify the TM mode. To simplify the discussion, we consider the so-called electrostatic limit, it in which the wavenumber $q$ of the TM mode is much larger than the free-space wavenumber $\frac{\omega}{c}$. This is well-satisfied in monolayer optical phonons in 2D. With this approximation, the condition for a TM mode corresponding to an optical phonon is
\begin{equation}
\epsilon_{\mathrm{env}} + \frac{1}{2}\epsilon_0 q \Pi(q,\omega) \equiv \epsilon_{\mathrm{RPA}}(\omega) = 0.
\end{equation}
The quantity $\epsilon_{\mathrm{RPA}}(\omega) = 0$ coincides with the phonon contribution to the longitudinal dielectric function of the monolayer in the random-phase approximation.

>>>>

% --- PACKAGES ---
\usepackage{graphicx}
\usepackage[centering,hmargin=20mm,tmargin=30mm,bmargin=25mm]{geometry}
\usepackage{multirow}
\usepackage{newtxtext}
\usepackage[cmintegrals]{newtxmath}

%-----SI UNITS-----
\usepackage{siunitx}
\sisetup{range-phrase =\text{\,--\,},
	list-units   =single,
	range-units  =single,
	list-pair-separator = {\ \text{and}\ },
	list-separator = {,\ \linebreak[0]},
	list-final-separator = {,\ \linebreak[0]\text{and}\ }}

%--- REFERENCING ---
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks,
	linkcolor={blue!75!black!80!yellow},
	citecolor={blue!75!black!80!yellow},
	urlcolor={blue!75!black!80!yellow}
}

%--- CAPTIONS (sans serif) ---
\makeatletter
\renewcommand\@make@capt@title[2]{%
	\@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
	\sffamily{\textbf{#1}}\@caption@fignum@sep#2
}%
\renewcommand\figurename{Figure}
\makeatother

% --- SPACING ---
\usepackage{microtype} % Kerning and symbol-stretching
\usepackage{xspace}

@@

% --- COMMANDS ---
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\sgn}{\operatorname{sgn}}

@@

%abbreviations
\newcommand{\ie}{i.e.\@\xspace} %Gobble-spaces of the "small" type (small is ensured by adding \@)
\newcommand{\cf}{cf.\@\xspace}
\newcommand{\eg}{e.g.\@\xspace}
\newcommand{\etc}{etc.\@\xspace}

@@

% ----- COMMENTS AND META-ANNOTATIONS -----
\usepackage{textcomp} % for \textrightarrow
\usepackage{xifthen}
\usepackage{etoolbox}
\newboolean{togglecomments}
\newboolean{togglechanges}

% toggle to true to see comments (otherwise hidden)
\setboolean{togglecomments}{true}  
% toggle to false to see mixed versions (otherwise edits are shown exclusively)
\setboolean{togglechanges}{false}

\newcommand{\comment}[2]{%
    \ifbool{togglecomments}%
    {\textcolor{blue!70!black}{\small\textsf{%
    \textsuperscript{\textsc{\textsf{\MakeLowercase{#1}}}}%
    [#2]}}} % if true, show comments
    {}}     % if false, do nothing
\newcommand{\swap}[2]{\ifbool{togglechanges}
    {#2}  % TC-only version
    {\textcolor{red!70!black}{[#1]}\text{\textrightarrow{}}\textcolor{green!50!black}{[#2]}}}
\newcommand{\remove}[1]{\ifbool{togglechanges}
    {}    % TC-only version
    {\textcolor{red!70!black}{#1}}}
\newcommand{\inset}[1]{\ifbool{togglechanges}
    {{}#1}  % TC-only version
    {\textcolor{green!50!black}{{}#1}}}
\makeatletter
\newcommand{\citeremind}[1]{%
	\unskip%
    \textcolor{blue!75!black!80!yellow}{${}^\blacksquare$%
	\ifthenelse{\isempty{#1}}{}{\textsuperscript{\tiny\textsf{#1}}}%
	}\xspace}
\makeatother
\newcommand{\FIXME}[1]{\textcolor{red!70!black}{[#1]}}
\newcommand{\optional}[1]{\textcolor{yellow!70!gray}{#1}}

% --- DOCUMENT ---
\begin{document}

% --- TITLE & AUTHORSHIP ---
\title{Nanophotonics with optical phonons in two dimensions}

\author{Nicholas Rivera}
\email{nrivera@seas.harvard.edu}
\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
%
\author{Thomas Christensen}
\affiliation{\MITPhy}
%
\author{Prineha Narang}
\email{prineha@seas.harvard.edu}
\affiliation{\HarvardSEAS}

@@

% --- ABSTRACT ---
\begin{abstract}
    Extreme confinement of electromagnetic energy by phonon-polaritons \swap{promises}{holds the promise of} extremely strong and novel forms of control over the dynamics of matter. 
    To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems. 
    \comment{tc}{It is not necessarily clear to me why 2D phonon-polaritonics would be an obvious component in bringing phonon-polaritons to their ``ultimate limit''?}
    Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.  
    Does this lack of LO--TO splitting imply the \swap{lack}{absence} of a \swap{strongly confined electromagnetic (phonon-polariton) mode}{phonon-polariton} in polar monolayers? 
    Here, we derive a universal form for the spatially-nonlocal dielectric function of a polar monolayer specified by: the TO phonon frequency, the non-zero group velocity of the LO phonon, and the phonon lifetime. 
    \comment{tc}{I'm a bit hesitant to highlight the ``spatially-nonlocal'' dielectric function: the real reason this (simplified) form is nonlocal is just dimensional---\eg graphene's dielectric function is also nonlocal in a similar fashion (even in the low-momentum, low-frequency limit). The associated surface conductivity is local; so the in-plane polarization response is also local.} 
    Our analysis reveals that the phonon-polariton of the bulk is \swap{``replaced''}{essentially supplanted/replaced} by the LO phonon\inset{ of the two-dimensional system}.
    We \swap{present}{discuss} the confinement and propagation losses of the LO \swap{phonons}{phonon-polaritons}, finding that high \remove{LO phonon} confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques. 
    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to \swap{2D phonon channels}{coupling with localized 2D phonon-polaritons}, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN \swap{phonons}{phonon-polaritons}.
\end{abstract}

@@

\comment{tc}{Both `phonon-polaritons' and `phonon polaritons' are used; let's settle on one.}

\comment{tc}{For title, how about `Phonon polaritons in two-dimensional hexagonal boron nitride'?}

% --- INTRODUCTION ---
Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, \swap{offer}{hold} great promise for deeply sub-diffractional control of electromagnetic fields at mid-\swap{IR}{infrared} and \swap{THz}{terahertz} frequencies. \swap{P}{Qualitatively, p}honon polaritons share many features in common with plasmon polaritons in conductors. \swap{In}{E.g., in} recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}. Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy,\citeremind{} radiative heat transfer~\cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters~\cite{kumar2015tunable,rivera2017making,kurman2018control}. 
In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which \swap{shrink}{reduce} the in-\remove{plane} and out-of-plane wavelength of polaritons monotonically with decreasing film thickness~\cite{dai2014tunable,dubrovkin2018ultra}. 
\comment{tc}{The previous and following sentences both start with `many cases' here; best to rewrite one (probably the latter).}
In many cases, a monolayer is the ultimate limit of this effect. Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions. 
\comment{tc}{I'm not sure that starting with `however' in this sentence is optimal: we want to convey an additional motivation for studying them (specifically, the transition from 3D to 2D is nontrivial)---maybe this sentence could be rewritten with this perspective in mind.}
However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO--TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point~\cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.

In this Letter, we develop \swap{core results in}{a framework and a several concrete examples for understanding the} the nascent field of optical phononics in two-dimensional materials\swap{ by deriving}{. We do so by deriving} a universal form for the dielectric function of a polar monolayer, \swap{incorporating spatially nonlocal effects}{including spatial nonlocality}, which \swap{is specified fully in terms of the dispersion relation of LO and TO phonons in two dimensions}{which depends solely on the LO and TO phonon frequencies---and their dispersion with momentum---in the two-dimensional system}.
Using parameters from Ref.~\citenum{sohier2017breakdown} for \swap{ important monolayers such as }{the canonical two-dimensional polar monolayer---}hexagonal boron nitride\swap{, }{---}we present the confinement and propagation losses of \swap{these modes}{the supported phonon-polaritons}, identifying the frequency region where they should be most easily detected. \swap{We then}{Finally, we} show that these modes \swap{essentially provide the ultimate limit of}{enable extreme} light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter--LO phonon coupling\swap{, which due to}{. Given} the narrow bandwidth of phonon-polariton resonances, \swap{can lead to }{this suggests} the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics \swap{for an external emitter coupled with optical phonons in hexagonal boron nitride}{in a two-dimensional hexagonal boron nitride platform}.

% -----------------
\section{Electrodynamics of optical phonons in two-dimensions}
\comment{tc}{Structurally, this section would benefit from being split into two sections: one solely tasked with deriving/establishing the forms of the (non-interacting) polarization-polarization response function---and possibly its equivalent conductivity; and another one solely tasked with deriving and discussing the dispersion relation for the associated longitudinal optical modes. Right now, this section goes back and forth between the two, which (I think) reduces its focus and gives way for unnecessary repetition.}

In this section, we develop \swap{the}{a} theory of electromagnetic \swap{waves associated with}{response due to} optical phonons in two\swap{ dimensions}{-dimensional systems}. The crux of the theory \swap{is determining}{involves} the dielectric \swap{permittivity}{function} associated with the monolayer. 
\swap{It is obtained by considering the response of the ions in the monolayer to a longitudinal electromagnetic wave whose electric field is the gradient of a scalar potential $\phi$.}{To obtain it, we consider the response of the ions of the monolayer due(/when subjected) to a longitudinal electromagnetic potential $\phi$}. The \swap{coupling}{interaction} Hamiltonian is \swap{then given by}{in this case simply}

@@

    H_{\mathrm{int}} = \int \dd^2x ~\rho \phi = -\int \dd^2x~ (\nabla\cdot\mathbf{P})\phi,
    \label{eq:coupling}

@@

with $\rho$ the \inset{induced} charge density and $\mathbf{P}$ the \inset{total} polarization density associated with the ionic motion. 
\swap{The polarization \inset{density} can be expanded in powers of the displacement of the ions. In particular, we can consider the polarization $P_i$ generated in the monolayer when atom $\kappa$ of the unit cell is displaced along the $j$ direction by amount $u_{\kappa,j}$. To lowest order:}{Within linear response, the induced polarization density can be straightforwardly evaluated from the displacement amplitude $\mathbf{u}_{\kappa}$ of every atom $\kappa$ within the unit cell. Specifically, to first-order, the induced polarization is:}
\comment{tc}{I don't see how the second equality can be consistent: since, by Taylor expansion, $P_i = P_i^0 + u_{\kappa,j}\partial_{u_{\kappa,j}}P_i + \mathcal{O}(u_{\kappa,j}^2)$---so where does the unit cell area $\Omega$ come from? The issue seems to arise from alternatingly interpreting the LHS/RHS $P_i$ as total/density quantities. The right-most equation also seems to be missing a $\kappa$-subscript in $u_{j}$.}

@@

where $Z_{\kappa,ij}$ is the Born effective charge tensor of ion $\kappa$ and $\Omega$ is the unit cell area.
\comment{tc}{We do not give a definition of $P_i^0$: we also do not comment on what the implications are if it is nonzero (it could be..!)---as far as I can see, we are assuming it is zero for the remaining analysis.}
With this relation between polarization and ionic displacements, \swap{we have expressed the Hamiltonian as a coupling between the scalar potential and ionic displacements}{the interaction Hamiltonian in Equation~\eqref{eq:coupling} couples the scalar potential and the ionic displacements}. 
We consider the response of the monolayer to a scalar potential of the form $\phi\inset{(\mathbf{r})} = \phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}$, where $\mathbf{q}$ is a two-dimensional wavevector in the plane of the monolayer. Such a potential corresponds to a longitudinal electric field \remove{of the form} $\mathbf{E}\inset{(\mathbf{r})} = \iu\mathbf{q}\swap{\phi(\mathbf{q},\omega)\e^{\iu\mathbf{q}\cdot\mathbf{r}-\iu\omega t}}{\phi(\mathbf{r})}$. 
\comment{tc}{I don't think it really ``follows'' from the preceding that $P_i = ... E_{j,\text{tot}}(\mathbf{q},\omega)$. First, this specializes to the RPA via a mean-field approximation, and second, it introduces a response function (so, really, it is more of a definition). It is probably better to phrase these next few sentences rather carefully and with specificity, \eg with references to the usual Kubo formula/approach etc. It may also make sense to split the $P_i = ... E_{j,\text{tot}}(\mathbf{q},\omega)$ off onto its own line, since this is the equation that lends practical utility to the $\boldsymbol{\Pi}$ response function. 
Finally, I don't think we actually need to mention the RPA or anything like it at this point: so far, this is just the non-interacting polarization-polarization response---we effectively arrive at the RPA later on, when we start doing electromagnetic calculations (\ie around Eq.~\eqref{eq:2dmaxwell}); only then do we actually make use of $\mathbf{P} \propto \boldsymbol{\Pi}\mathbf{E}_{\mathrm{tot}}$.}
It follows that the induced polarization takes the form $P_i\inset{(\mathbf{q})}  = \epsilon_0\Pi_{ij}(\mathbf{q},\omega)E_{j,\mathrm{tot}}(\mathbf{q},\omega)$, where $\Pi_{ij}(\mathbf{q},\omega)$ is the polarization-polarization response function of the monolayer and  $E_{j,\mathrm{tot}}(\mathbf{q},\omega)$ is taken as the \textit{total} field which is a sum of the applied field and the field created by induced polarization. The polarization-polarization response function is given by

@@

    \boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\Big(\e^{-\beta E_m}-\e^{-\beta E_n} \Big),

@@

where $m,n$ are states of the phononic Fock space of the monolayer, $P_{mn,i} = \sum_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes,
\comment{tc}{Probably best to include $\mathbf{q}$ dependence explicitly in both LHS and RHS of the definition of $P_{mn,i}\textcolor{red!50!black}{(\mathbf{q})}$.}
$E_{m}$ ($E_n$) is the energy of state $m$ ($n$), $\beta\inset{\equiv 1/k_{\text{\textsc{b}}}T}$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function.
Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by: 
\comment{tc}{It's not clear how we obtain Eq.~\eqref{eq:polpolresponse} from Eq.~\eqref{eq:2dsusceptibility}: \eg, I'm not sure what you mean by the harmonic approximation here? There's also a low-temperature approximation at play here. Finally, every phonon band except the LO one is dropped. Finally, we have not previously defined the $\mathbf{P}(\mathbf{q})$ polarization operator. I guess I'm just asking for us to be more explicit and detailed here.}

@@

    \Pi_{qq}(\mathbf{q},\omega) = \frac{1}{\epsilon_0\hbar\Omega} \frac{2\omega_{\mathbf{q},\mathrm{L}}}{\omega^2_{\mathbf{q},\mathrm{L}}-\omega^2}|\hat{\mathbf{q}}\cdot\langle 1_{\mathbf{q},\mathrm{L}}|\mathbf{P}(\mathbf{q})|0_{\mathbf{q}\sigma}\rangle|^2,
    \label{eq:polpolresponse}

@@

where L-subscripts denote longitudinal polarization, $|0_{\mathbf{q},\mathrm{L}}\rangle$ ($|1_{\mathbf{q},\mathrm{L}}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$. The frequency $\omega_{\mathbf{q},\mathrm{L}}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO--TO splitting.\citeremind{}
This is consistent with the fact that LO--TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility. Coulomb interactions are accounted for in the random phase approximation, 
\comment{tc}{Another reason for us to be more clear about the implications of the mean-field assumption before Eq.~\eqref{eq:2dsusceptibility}; might as well establish equivalence with RPA clearly and only once.}
and to include them in the single-particle response \swap{is an}{would amount to an} uncontrolled double-counting. 
\comment{tc}{Could it be better to interleave this $qq$-comment into the lead-up sentence for Eq.~\eqref{eq:polpolresponse}? Basically, we want to say that we specialize to the diagonal components. Finally; is it really necessary for us to make that scope reduction here? Wouldn't the formula nearly be simpler without it?}
The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  
\comment{tc}{It might be better to break this off into a new paragraph here (and possibly merge (and maybe shorten) with the following paragraph) since this is sort of orthogonal to the previous stuff.}

@@

\comment{tc}{I don't think this needs to be as long as it currently is---also, some of the details, \eg position of Reststrahlen band, assumption of long wave-length etc., would probably fit in more meaningfully when those approximations are actually applied. For Eq.~\eqref{eq:2dmaxwell}, only the homogeneity assumption is necessary (corresponding to in-plane translation invariance)}
To strip the analysis to its bare essentials, we consider optical phonon response with in-plane isotropy in the long-wavelength limit arising from in-plane LO oscillations. A relevant example of a system where these conditions are satisfied is in a hexagonal boron nitride monolayer, considering optical phonons in the so-called upper Reststrahlen band which in bulk spans the frequency range of \SIrange[range-phrase=\text{ to }]{1360}{1610}{\per\cm}. In a monolayer geometry with translation invariance and in-plane isotropy, the electromagnetic modes of Maxwell's equations can be composed into transverse magnetic (TM) and transverse electric (TE), where the magnetic or electric field respectively is transverse to the in-plane wavevector of the mode. In practice, it is the TM mode which is associated with highly confined electromagnetic waves that are of use in nanophotonics. TE waves are not supported at the same frequency as transverse magnetic waves, and require exotic conditions to be realized in two-dimensional materials. Given the isotropy, we may suppress indices from the response tensor. We may also consider without \remove{further} loss of generality a TM mode with wavenumber $q$ along the $x$-direction in the monolayer and magnetic field $H(z)\e^{\iu qx-\iu\omega t}$ along the $y$-direction of the monolayer. The direction transverse to the monolayer is denoted as $z$.  With these definitions in place, the Maxwell equation satisfied by the magnetic field is 
\begin{equation}\label{eq:2dmaxwell}
    \bigg(-\frac{\dd^2}{\dd{}z^2}+q^2-\epsilon_{\mathrm{env}}\frac{\omega^2}{c^2} \bigg)H(z) = 0.
\comment{tc}{I'm not sure this needs to be its own equation. Even if it does, I don't think that the term $\epsilon_{\mathrm{env}}\omega^2/c^2$ is correct; shouldn't this really be $\epsilon_{1,2}\omega^2/c^2$, with choice of $\{1,2\}$ depending on sign of $z$?}
We consider a solution of the form $H(z) = \e^{-\kappa_1 z}$ for $z > 0$ with $\kappa_1 = \sqrt{q^2-\epsilon_1\frac{\omega^2}{c^2}}$ and a solution of the form $H(z) = c\e^{\kappa_2 z}$ for $z < 0$ with $\kappa_2 = \sqrt{q^2-\epsilon_2\frac{\omega^2}{c^2}}$. 
\comment{tc}{Should probably choose a different name for the constant `$c$' to avoid clashing with the speed of light $c$. Maybe $c_1$ or $a$. Or, alternatively, let $H(z) = h_\pm\e^{\mp\kappa_\pm z}$ with $\kappa_\pm = \sqrt{q^2-\epsilon_\pm k_0^2}$ with upper/lower signs applicable for $\pm z>0$ (instead of using $\{1,2\}$ subscripts)---and then specify the boundary condition as $h_-/h_+ = \ldots (=c)$. Personally, I favor the latter option but it requires a few more fixes. If you go with that option, you can probably drop the parentheses around $H_y^{(\pm)}$ in the following.}
The boundary condition on the magnetic field is $H_y^{(+)}-H_y^{(-)} = -K_x = \iu\omega\mathbf{P}_x$ where $\mathbf{K}$ is the surface current density, expressed through the polarization density $\mathbf{P}$ of the monolayer. 
This condition enforces $c =(1+\frac{\epsilon_0}{\epsilon_2} \kappa_2\Pi)^{-1}$. 
Continuity of the electric field in the $x$ direction enforces $\frac{\epsilon_2}{\epsilon_1} = -\frac{\kappa_2}{\kappa_1}c$. 
\swap{These two conditions fully specify the TM mode}{Combining the two conditions, we obtain the usual dispersion equation for the TM mode of a polarizable 2D monolayer, namely $\epsilon_1/\kappa_1 + \epsilon_2/\kappa_2 +\epsilon_0\Pi=0$.}. 
\swap{To simplify the discussion, we consider the so-called electrostatic limit, it in which the wavenumber $q$ of the TM mode is much larger than the free-space wavenumber $\frac{\omega}{c}$. This is well-satisfied in monolayer optical phonons in 2D. With this approximation, the condition for a TM mode corresponding to an optical phonon is}
